Semi-Supervised Factored Logistic Regression for High-Dimensional Neuroimaging Data

### 2015-06-04 paper_nips/paper.tex
results: added notice on reward/punishment cognitive tasks

\paragraph{Individual effects of dimensionality reduction and rest data.}
We first quantified the gain of introducing a bottleneck layer
disregarding the autoencoder.
To this end, ordinary logstic regression was juxtaposed
with FLogReg  at $\lambda=1$ (no autoencoder).
For this experiment, we increased the difficulty of the classification problem
by including data from all 38 HCP tasks.
Indeed, increased class separability in component space,
as compared to voxel space, entails differences in
generalization performance of
$\approx{17\%}$ (Figure \ref{fig_dimred}).
\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/accuracies.pdf}
\end{centering}
\vspace{-0.5cm}
\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$).
Multinomial logistic regression operating in voxel space (\textit{blue bars})
was compared to 
factored logistic regression operating in $20$
(\textit{left plot}) and $100$ (\textit{right plot})
latent modes (\textit{grey bars}).
Autoencoder or rest data were not used for these analyses
($\lambda=1$).
Ordinary logistic regression yielded 77,7\% accuracy out of sample, while
factored logistic regreesion scored at 94,4\% ($n=20$) and 94,2\% ($n=100$).
Hence, compressing the voxel data into a component space for classification
achieves higher task separability.
}
\label{fig_dimred}
\end{figure}

@@

\section{Discussion and Conclusion}
% Conceptual recap+gain of current paper
Using the flexibility of autoencoders, we
learn the optimal decomposition from high-dimensional
voxel brain space into the most important activity patterns for a
supervised learning question.
%
From a machine-learning perspective,
factorization of the logistic regression weights
can be viewed as transforming a ``multi-class problem'' into
``multi-task learning''.
%
The higher generalization accuracy and feature recovery, comparing to
ordinary logistic regression, hold potential
for adoption in various neuroimaging analyses.
Besides increased performance, these models are more interpretable by
automatically learning a mapping to and from a brain-network space.
%
This domain-specifc classification algorithms
encourages departure from the artificial and statistically
less attractive voxel space.
Neurobiologically,
brain activity underlying defined mental operations
can be explained by linear combinations of the main activity
patterns.
%
That is,
fMRI data probably concentrate near
a low-dimensional manifold of
characteristic brain networks combinations.
Extracting fundamental building blocks of brain organization might
facilitate the quest for the cognitive primitives of
human thought.
%
We hope that these first steps stimulate development towards
powerful semi-supervised classification methods in systems neuroscience.

>>>>

\paragraph{Individual effects of dimensionality reduction and rest data.}
We first quantified the gain of introducing a bottleneck layer
disregarding the autoencoder.
To this end, ordinary logstic regression was juxtaposed
with FLogReg  at $\lambda=1$ (no autoencoder).
For this experiment, we increased the difficulty of the classification problem
by including data from all 38 HCP tasks.
Indeed, increased class separability in component space,
as compared to voxel space, entails differences in
generalization performance of
$\approx{17\%}$ (Figure \ref{fig_dimred}).
%
Notably,
the cognitive tasks on reward and punishment processing
are among the least predicted with ordinary but well predicted with
factored logistic regression
(tasks 3 and 4 in Figure \ref{fig_dimred}).
These experimental conditions have indeed been reported to exhibit
highly similar neural activity patterns in
GLM analyses of that dataset \cite{barch2013}.
Consequently, also local activity differences
(in the striatum and visual cortex in this case) can
be successfully captured by brain-network modelling.
%
\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/accuracies.pdf}
\end{centering}
\vspace{-0.5cm}
\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$)
sorted from lowest to highest.
Multinomial logistic regression operating in voxel space (\textit{blue bars})
was compared to 
factored logistic regression operating in $20$
(\textit{left plot}) and $100$ (\textit{right plot})
latent modes (\textit{grey bars}).
Autoencoder or rest data were not used for these analyses
($\lambda=1$).
Ordinary logistic regression yielded 77,7\% accuracy out of sample, while
factored logistic regreesion scored at 94,4\% ($n=20$) and 94,2\% ($n=100$).
Hence, compressing the voxel data into a component space for classification
achieves higher task separability.
}
\label{fig_dimred}
\end{figure}

@@

\section{Discussion and Conclusion}
% Conceptual recap+gain of current paper
Using the flexibility of factored models,
we learn the low-dimensional representation from high-dimensional
voxel brain space that is most important for
prediction of cognitive task sets.
%
From a machine-learning perspective,
factorization of the logistic regression weights
can be viewed as transforming a
``multi-class classification problem''
into a ``multi-task learning problem''.
%
The higher generalization accuracy and feature recovery, comparing to
ordinary logistic regression, hold potential
for adoption in various neuroimaging analyses.
Besides increased performance, these models are more interpretable by
automatically learning a mapping to and from a brain-network space.
%
This domain-specifc learning algorithm
encourages departure from the artificial and statistically
less attractive voxel space.
Neurobiologically,
brain activity underlying defined mental operations
can be explained by linear combinations of the main activity
patterns.
%
That is,
fMRI data probably concentrate near
a low-dimensional manifold of
characteristic brain networks combinations.
Extracting fundamental building blocks of brain organization might
facilitate the quest for the cognitive primitives of
human thought.
%
We hope that these first steps stimulate development towards
powerful semi-supervised representation extraction
in systems neuroscience.
